# Naive Bayes Classifier
Naive Bayes is a supervised model that will allow us to 
classify the texts into the two categories. It is a 
probabilistic classifier that uses Bayes’ theorem to 
predict the class of a given text. The theorem states 
that the probability of an event A occurring given that 
event B has occurred is equal to the probability of event 
B occurring given that event A has occurred, multiplied by 
the probability of event A occurring, divided by the 
probability of event B occurring. In the context of this 
project, the probability of a text belonging to a given  
category given that it contains a given word is equal to 
the probability of the word occurring in the text given. 

## Imports
```{r}
require(quanteda)
require(quanteda.textmodels)
require(dplyr)
require(caret)
```

## Data
```{r}
# Check datatype of the data
df <- read.csv("dataset/drugs_cleaned.csv")

sample_reviews <- read.csv("dataset/sample_reviews.csv")
sample_review <- sample_reviews$review[0]
```

## Preprocessing steps
1. Tokenization: Breaking the text data into individual words or phrases.
2. Lowercasing: Converting all text to lowercase to reduce the dimensionality of the data.
3. Removing Stop words: Removing commonly used words such as "the", "and", "is", etc. which do not provide much information.
4. Stemming/Lemmatization: Reducing words to their base form (e.g., "running" to "run") to reduce dimensionality.
5. Removing punctuation and special characters.
6. Removing numbers
7. Removing HTML tags and URLs
8. Creating a bag of words representation of the text data.

```{r}
to_dfm <- function(s_corpus) {
    # 1. Tokenization
    s_tokens <- tokens(s_corpus)

    # 2. Lowercasing
    s_tokens <- tokens_tolower(s_tokens)

    # 3. Removing Stop words except negations
    stop_keep <- c("will not", "ok", "neither", "won't", "cannot", "no", "nor", "not")
    stop_w <- stopwords("english") %>% setdiff(stop_keep)
    s_tokens <- tokens_remove(s_tokens, stop_w)

    # 4. Stemming
    stem_tokens <- tokens_wordstem(
        s_tokens,
        language = quanteda_options("language_stemmer")
    )

    # 5. Removing punctuation and special characters
    stem_tokens <- tokens_remove(stem_tokens, pattern = "[[:punct:]]")

    # 6. Removing numbers
    stem_tokens <- tokens_remove(stem_tokens, pattern = "[[:digit:]]")

    # 7. Removing HTML tags and URLs
    stem_tokens <- tokens_remove(stem_tokens, pattern = "<[^>]+>")

    # 8. Creating a bag of words representation of the text data.
    s_dfm <- dfm(stem_tokens)

    return(s_dfm)
}

# Create corpus
gen_corpus <- function(df, text_field) {
    cor <- corpus(df, text_field = text_field)
    # Add id to corpus
    df$id <- seq_len(nrow(df))
    cor$id <- df$id

    return(cor)
}
```

## Format and split data 
```{r}
set.seed(123)

# Create corpus
cor <- gen_corpus(df, "review")
dfm <- to_dfm(cor)

# Split test and train data
train_index <- createDataPartition(df$rating_type, p = 0.8, list = FALSE)

# Split test and train data
train_dfm <- dfm_subset(dfm, id %in% train_index)
test_dfm <- dfm_subset(dfm, !id %in% train_index)
```

## Train Model
```{r}
# Train model
nb_model <- textmodel_nb(train_dfm, train_dfm$rating_type)
summary(nb_model)
```

## Predict
```{r}
# Makes sure all features are present in the test data
matched_dfm <- dfm_match(test_dfm, features = featnames(train_dfm))

actual_class <- matched_dfm$rating_type
predicted_class <- predict(nb_model, matched_dfm)
tab_class <- table(actual_class, predicted_class)
tab_class
```

```{r}
# Confusion matrix
confusionMatrix(tab_class)
```